Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-06-27 analysis/memo2_TAS_reviewers.Rmd
minor edits to responses to Reviewer 2

Thank you for suggesting this contrast between _Batson_ law and these State laws. We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." (@flowers2019, p. 2241) We did not use the phrase "significant factor," because it does not appear in _Miller-El v. Dretke_, 545 U.S. 231 (2005). At best, the Court there wrote: "Comparing his strike with the treatment of panel members who expressed similar views supports a conclusion that race was significant in determining who was challenged and who was not. _Id._ at 252 (footnote omitted).

@@

Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal strike bias (the credible interval includes zero), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal strike bias (the credible interval excludes zero). 
We have revised the Discussion to further address high incompatibility. As we observe there, because no one can directly observe $b_{curr}$ and $b_{hist}$, they cannot know for sure whether their case involves high incompatibility. Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility. As for what a judge might do in this situation, a judge might not find a _Batson_ prima facie case for that reason. Or the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred. In the reverse situation of high incompatibility, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but (erroneously) does not detect illegal bias when $\alpha > 0$. Here, we agree that a judge should worry about erroneously failing to detect strike bias due to high incompatibility. Down-weighting the historical data here increases the support for a finding that illegal strike bias likely occurred.

>>>>

Thank you for suggesting this contrast between _Batson_ law and these State laws. We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." _Flowers v. Mississippi_, 139 S. Ct. 2228, 2241 (2019). We did not use the phrase "significant factor," because it does not appear in _Miller-El v. Dretke_, 545 U.S. 231 (2005). At best, the Court there wrote: "Comparing his strike with the treatment of panel members who expressed similar views supports a conclusion that race was significant in determining who was challenged and who was not. _Id._ at 252 (footnote omitted).

@@

Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal strike bias (the credible interval includes zero), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal strike bias (the credible interval excludes zero).

We have revised the Discussion to further address high incompatibility. As we observe there, because no one can directly observe $b_{curr}$ and $b_{hist}$, they cannot know for sure whether their case involves high incompatibility. Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $\alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility.

A judge might respond to this situation, in two ways.  First, a judge might not find a _Batson_ prima facie case for that reason. Second, the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred.

In the reverse situation of high incompatibility, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but (erroneously) does not detect illegal bias when $\alpha > 0$. Here, we agree that a judge should worry about erroneously failing to detect strike bias due to high incompatibility. Down-weighting the historical data here, however, increases the support for a finding that illegal strike bias likely occurred, all else equal.
